DateLife: leveraging databases and analytical tools to reveal the dated Tree of Life

### 2024-02-15 data-raw/felis-chronogram-plot.Rmd
Edits from final decision

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
getwd()
source(file = "R/plot_node_ages_v2.3.R")
library(ape)
library(datelife)
```

@@

phylo_all <- datelife::datelife_search(input = my_datelife_query,
                                           summary_format = "phylo_all")
class(phylo_all)
ls(phylo_all)
length(phylo_all)
```

@@

```{r}
lapply(phylo_all, ape::branching.times)

plot_max_time <- 846
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
getwd()
source(file = "../R/plot_node_ages_v2.3.R")
library(ape)
# detach("package:datelife", unload=TRUE)
devtools::install(pkg = "../../lunasare/datelife/.")

# devtools::install_github(repo = "phylotastic/datelife", ref = "ott-names")
library(datelife)
```

@@

phylo_all <- datelife::datelife_search(input = my_datelife_query,
                                           summary_format = "phylo_all")
class(phylo_all)
ls(phylo_all)
length(phylo_all)
```
### Visualizing all source chronograms

```{r}
plot_max_time <- 845
source_chronogram_file_names <- paste0("felis-source-chronogram-", 1:length(phylo_all))
datelifeplot::plot_phylo_all(chronograms = phylo_all,
                             folder_name = paste0("felis-source_chronograms"),
                             file_name = source_chronogram_file_names,
                             plot_height = 600,
                             plot_width = 600,
                             mai1 = 1,
                             mai2 = 0,
                             mai3 = 1,
                             mai4 = 5,
                             omi1 = 1,
                             omi2 = 0,
                             omi3 = 0,
                             omi4 = 0,
                             write = "pdf",
                             plot_type = "phyloch",
                             max_depth = plot_max_time,
                             edge.width = 2,
                             cex_tiplabels = 1,
                             cex_axislabel = 1.5,
                             cex_axis = 2,
                             cex_title = 2.5,
                             pos_title = -2.8,
                             pos_axis = 2,
                             axis_label = "")
```

@@

```{r}
lapply(phylo_all, ape::branching.times)
class(phylo_all[[9]]) <- "phylo"
datelifeplot::plot_chronogram(chronogram = plot(phylo_all[[9]]))
```

### 2024-02-15 manuscript/manuscript.Rmd
Edits from final decision

At the time of writing of this manuscript
(`r format(Sys.time(), "%b %d, %Y")`),
DateLife's chronogram database latest version consist of `r format(length(opentree_chronograms$trees), big.mark=",")` chronograms published in `r length(unique(opentree_chronograms$studies))` different studies, encompassing close to 100 k (99 474) different species. It is curated from OpenTree's phylogenetic database, the Phylesystem, an open database of expert and peer-reviewed phylogenetic knowledge with rich metadata and a wide taxonomic scope [@mctavish2015phylesystem]. We expect DateLife's database to largely overlap with OpenTree's phylogenetic database taxonomic coverage, where Chordata and Embryophyta are nearly fully sampled. In contrast, Bacteria, Fungi, Nematoda, and Insecta, currently present a large gap between the number of named species and what has phylogenetic information in OpenTree’s synthetic tree. It is likely that users working with the former groups will get results from a DateLife analysis. If none of the user’s species are found, the software will indicate the lack of age data for the queried taxa in the database.

>>>>

At the time of writing of this manuscript
(`r format(Sys.time(), "%b %d, %Y")`),
DateLife's chronogram database latest version consist of `r format(length(opentree_chronograms$trees), big.mark=",")` chronograms published in `r length(unique(opentree_chronograms$studies))` studies, encompassing 99 474 species. It is curated from OpenTree's phylogenetic database, the Phylesystem, an open database of expert and peer-reviewed phylogenetic knowledge with rich metadata and a wide taxonomic scope [@mctavish2015phylesystem]. We expect DateLife's database to largely overlap with OpenTree's phylogenetic database taxonomic coverage, where Chordata and Embryophyta are nearly fully sampled. In contrast, Bacteria, Fungi, Nematoda, and Insecta, currently present a large gap between the number of named species and what has phylogenetic information in OpenTree’s synthetic tree. It is likely that users working with the former groups will get results from a DateLife analysis. If none of the user’s species are found, the software will indicate the lack of age data for the queried taxa in the database.
